---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# streetnamer

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

The goal of `streetnamer` is to facilitate the matching of street name to their Wikidata identifiers.

This is a pre-release version. Some elements of the interface work as expected, but the package it's unusable in its present state. You have been warned. 

## Installation

You can install the development version of `streetnamer` with:

``` r
remotes::install_github("giocomai/latlon2map") # required dependency not on CRAN
remotes::install_github("EDJNet/tidywikidatar") # available on CRAN, but some features used by `streetnamer` may be added to the development version
remotes::install_github("giocomai/streetnamer")
```

This package relies heavily on [`tidywikidatar`](https://edjnet.github.io/tidywikidatar).

Since all three packages (`streetnamer`, `latlon2map`, and `tidywikidatar`) are being developed concurrently, leaving to each a separate group of tasks, at this stage updates impacting the app may occur to any of them. Hence, if anything is not working as expected, you are invited to update those packages before reporting.

## How does it work?

At this stage, not much really works. 

In order to get a preview of how the interface looks like, you can try running the following code chunks. 

Keep in mind that OpenStreetMap data for the whole country are downloaded when you first select a city, so be prepared to wait many minutes. Municipality-level data are cached and retrieved efficiently afterwards.

```{r testing app, eval = TRUE}
library("streetnamer")
library("latlon2map")
library("tidywikidatar")
options(timeout = 60000) # big timeout, as big downloads needed 

ll_set_folder(path = fs::path(fs::path_home_r(),
                              "R",
                              "ll_data"))
sn_set_data_folder(fs::path(fs::path_home_r(),
                            "R",
                            "sn_data"))

# tidywikidatar cache
tw_set_cache_folder(path = fs::path(fs::path_home_r(),
                            "R",
                            "tw_data"))

## or in a temporary folder for testing

# tw_set_cache_folder(path = fs::path(tempdir(),
#                                     stringi::stri_rand_strings(n = 1, length = 24)))
#             

tw_create_cache_folder(ask = FALSE)

## if using rstudio, I'd suggest you set open this in your default browser
## rather than in rstudio's enabling the following option

# options(shiny.launch.browser = .rs.invokeShinyWindowExternal)

# sn_run_app()
```

## Function naming conventions

`streetnamer` has two main types of functions:

- a set of functions used to facilitate processing, that can conventionally be used from the command line, or internally by the Shiny app: they all start with `sn_` followed by a verb, e.g. `sn_get_lau_street_names()`
- a set of functions that are in effect Shiny modules (see below). They typically start with `mod_sn_` and are currently not exported (as is customary for non-exported functions, they can be used with the triple `:`, e.g. `streetnamer:::mod_sn_street_info_app`) .


## Shiny modules

In order to facilitate development, as well as to allow integration of component parts of this app in spin-off projects, key components of the Shiny app have been developed as modules and can be tested independently.


### Module that shows info about Wikidata

```{r eval=FALSE}
streetnamer:::mod_sn_street_info_app(street_name = "Belvedere San Francesco",
                                     gisco_id = "IT_022205")
```


### Module for showing

### Module for exporting data

## What happens in the background

The selectors on the top and the left allow to pick a municipality, and then a street. 



When you click on a street name, a set of options to add data on a given street name appears. This is the choices that appear with this module:

```{r eval=FALSE}
streetnamer:::mod_sn_street_info_app(street_name = "Belvedere San Francesco",
                                     gisco_id = "IT_022205")
```

All the choices made in this interface are transformed into a data frame, that is written into a database:

```{r}
sn_write_street_name_wikidata_id(
  gisco_id = "IT_022205",
  country = "IT",
  street_name = "Belvedere San Francesco",
  person = TRUE,
  wikidata_id = "Q676555",
  gender = "male",
  category = "religion",
  tag = "",
  checked = TRUE,
  session = "testing",
  append = TRUE,
  overwrite = FALSE,
  disconnect_db = TRUE
)


street_info_df <- sn_get_street_name_wikidata_id(
  gisco_id = "IT_022205",
  street_name = "Belvedere San Francesco",
  country = "IT"
)

street_info_df %>% 
  dplyr::distinct(gisco_id, .keep_all = TRUE) %>% 
  tidyr::pivot_longer(cols = dplyr::everything(),
                      names_to = "type",
                      values_to = "value",
                      values_drop_na = FALSE,
                      values_transform = as.character) %>% 
  print(n = 100)
```

Each time the "confirm" button is clicked, a new row is added to the database. Hence, when you process the data you need to decided which criteria to use for keeping data, e.g. the most recent row, or the most confirmed. 

This set of data support a number of special cases, and different degrees of information that can be shared:

Done:
- data is confirmed at the country or city level - we expect data to be valid if confirmed at country levels, but especially with common surnames (or e.g. common names of saints, where one city has places dedicated to a locally born but globally less famous saint) it may be useful to check data at the city level
  - when checking if a street is tagged, this can be effectively done by filter for either the `gisco_id` column or the `country` column
- it is possible to ignore a given street name - in OpenStreetMap is relatively common to have some streets that do not have a proper street name, mostly because they are improperly tagged (e.g. just a number, or a hyphen), or because they have descriptive names that are not actually street names (e.g. "access ramp to hospital"). These should simply be ignored and not added in the count of total streets. 
  - this is expressed via the `ignore` column, with expected values either 1 (TRUE) or 0 (FALSE)
- make it possible to confirm that a street name is not named after a human, without adding anything else - this is useful because in some use cases the main point of interest is humans, and requiring to add a Wikidata identifier would needlessly prolong the checking times
  - this is expressed via the `person` column, with expected values either 1 (TRUE) or 0 (FALSE)
- make it possible to claim that a street is named after more than one person/individual
  - this is achieved by having a column with how many entities the street is dedicated to, `dedicated_to_n`. When reading the data, if `dedicated_to_n` is more than 1, then more than one row with data is expected to be found. Is is the responsibility of those who read the data do deal with potential inconsistencies

To do:


## Deduplication

- add Wikidata identifier of the actual street - this can be useful, as a number of properties are associated to it, possibly including different values for "named after" with qualifiers when street names changed
  - this is achieved with a separate column, `wikidata_street_id`. This should always be considered in combination with a given municipality. 


## Caching

Rather than adopting a separate caching infrastrucutre, `streetnamer` relies on the caching infrastructure of `tidywikidatar`. In brief, it generates separate tables with non-conflicting names in the same database used by `tidywikidatar` (be it a local SQLite or another odbc-compliant servers such as SQL)


## Deployed shiny app

Given that Shiny Server limits access to environment variables, for the deployed app a connection must be directly passed to `sn_run_app()`, and cannot be simply be set before startup (which works fine when running the app locally).


## Data sources

- OpenStreetMap data (© OpenStreetMap contributors) as kindly made available by [Geofabrik](http://download.geofabrik.de/)


## Desired features

It should be possible deal with the following circumstances:

- streets that are on OSM
- streets that are available on other lists, but not on OSM
- streets with wikidata id or without
- streets that are a person or not a person
- different streets that are the same street (deduplication)
- not a street / irrelevant
- single street has more wikidata id (e.g. dedicated to two individuals)
- add a tag for each street (maybe, free tag from a controlled vocabulary, e.g. to mark streets related to some issue that would not appear from relevant Wikidata identifier)

## Checks for things that should work




## On naming things

OpenStreetMap groups all sorts of roads, streets, squares, and paths under the confusing label of "highway". Within this package, the generic word used in function and documentation will be "streets", as the package is expected to be used chiefly in reference to urban centres.

## Contributing

Suggestions and contributions are welcome; they can be discussed via GitHub issues. 

## Copyright and credits

This package has been created by [Giorgio Comai](https://giorgiocomai.eu), data analyst and researcher at [OBCT/CCI](https://balcanicaucaso.org/), within the scope of [EDJNet](https://europeandatajournalism.eu/), the European Data Journalism Network.

It is distributed under the MIT license. 
